---
title: "SWC/DC Workshops in Canada"
output: rmarkdown::github_document
---

## Motivation

My goal in joining the Carpentries mentorship group is to help in community building in Canada.  In particular, I would like to increase both geographical and institutional diversity of Software Carpentry and Data Carpentry workshops.  I thought it might be useful to look at the locations of past workshops.

## Preamble

Load libraries and define paths.

```{r}
# for data frames
library(dplyr)

# for making maps
library(maptools)
library(rgdal)
library(raster)
library(maps)
library(mapdata)
library(ggmap)
library(marmap)
library(lattice)
library(RColorBrewer)
```

```{r}
pathData <- file.path('..', 'data');
```

## Data setup

The first step is to parse the html files for the Canadian locations.  This is kind of embarrassing... I really have to learn more regex!  

The following command partially parses the information from Canadian SWC workshops:

```{shell eval=FALSE}
grep -A1 "ca.png" ../data/2018-04-17-swc-list.html | grep -P "(20[0-9]{2}-[0-9]{2}-[0-9]{2})" | grep -oP "(20*).*" > ../data/2018-04-19-swc-parsed1.txt
```

Sanity check:
```{shell eval=FALSE}
grep -A1 "ca.png" ../data/2018-04-17-swc-list.html | grep -P "(20[0-9]{2}-[0-9]{2}-[0-9]{2})" | wc -l
```
There should be 125 workshops.

```{r}
# read in the partially parsed file
temp <- read.csv(
	file = file.path(pathData, '2018-04-19-swc-parsed1.txt'),
	header = FALSE,
	stringsAsFactors = FALSE,
	quote="",
	sep="/"
);

# sanity checks
print(nrow(temp))
print(head(temp))

# collect all the data in a dataframe 
swc <- data.frame(matrix(, nrow=nrow(temp), ncol=0))
swc$code <- tolower(substring(temp$V1, 12))
swc$year <- substr(temp$V1, 1, 4)
swc$month <- substr(temp$V1, 6,7)
swc$day <- substr(temp$V1, 9,10)
swc$location <- gsub("<", "", gsub('">',"", temp$V2))

# clean up some of the redundancies in "code" field
# done manually (hence boring code)
swc$simpleCode <- swc$code;
swc$simpleCode <- gsub("swc-", "", swc$simpleCode)
swc$simpleCode <- gsub("-python", "", swc$simpleCode)
swc$simpleCode[26] <- "uvic";
swc$simpleCode <- gsub("york-university", "yorku", swc$simpleCode)
swc$simpleCode <- gsub("lsiubc", "ubc", swc$simpleCode)
swc$simpleCode <- gsub("^waterloo", "uwaterloo", swc$simpleCode)
swc$simpleCode <- gsub("^alberta", "ualberta", swc$simpleCode)
swc$simpleCode <- gsub("^westernu", "uwo", swc$simpleCode)
swc$simpleCode <- gsub("smu", "saintmarys", swc$simpleCode)
swc$simpleCode[grep("University of Toronto", swc$location)] <- "utoronto";
swc$simpleCode[grep("University of Calgary", swc$location)] <- "ucalgary";
swc$simpleCode[grep("Queen", swc$location)] <- "queensu";
```

```{r}
# include province information
# again, also done manually
swc$province <- NA;
swc$province[swc$simpleCode %in% c("ubc", "ubco", "sfu", "uvic")] <- "British Columbia";
swc$province[swc$simpleCode %in% c("queensu", "utoronto", "toronto", "yorku", "mcmaster", "uwaterloo", "uoit", "brocku", "camh", "conestoga", "uguelph", "uwo", "sickkids", "wise-toronto", "oicr-toronto", "mozilla")] <- "Ontario";
swc$province[swc$simpleCode %in% c("mcgill", "uqam", "universite-laval", "umontreal", "montreal", "fmv")] <- "Quebec";
swc$province[grep("pycon", swc$simpleCode)] <- "Quebec";
swc$province[swc$simpleCode %in% c("mun")] <- "Newfoundland and Labrador";
swc$province[swc$simpleCode %in% c("halifax", "dalhousie", "acadia", "saintmarys")] <- "Nova Scotia";
swc$province[swc$simpleCode %in% c("ualberta", "ucalgary")] <- "Alberta";
swc$province[swc$simpleCode %in% c("umanitoba")] <- "Manitoba";
swc$province[swc$simpleCode %in% c("unb")] <- "New Brunswick";
swc$province[swc$simpleCode %in% c("caims")] <- "Saskatchewan";
```

```{r}
# find the number of SWC workshops per province
provinceCounts <- swc %>% group_by(province) %>% summarise(n=n()) %>% ungroup() %>% arrange(-n)
print(provinceCounts)
```

```{r fig.width=20}
# create colour gradient
colfunc <- colorRampPalette(c("yellow","red"))
colVals <- colfunc(max(provinceCounts$n));
colInd <- provinceCounts$n;
 
# get the map of Canada
Canada <- getData('GADM', country="CAN", level=1)

# match the province names in the table with the
provInd <- match(provinceCounts$province, Canada$NAME_1);
provInd[is.na(provInd)] <- 13;

# Very pretty code from: https://stackoverflow.com/questions/10763421/r-creating-a-map-of-selected-canadian-provinces-and-u-s-states?utm_medium=organic&utm_source=google_rich_qa&utm_campaign=google_rich_qa

# Specify a geographic extent for the map
# by defining the top-left and bottom-right geographic coordinates
mapExtent <- rbind(c(-156, 80), c(-68, 19))

# Specify the required projection using a proj4 string
# Use http://www.spatialreference.org/ to find the required string
# Polyconic for North America
newProj <- CRS("+proj=poly +lat_0=0 +lon_0=-100 +x_0=0 
            +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")

## Project the map extent (first need to specify that it is longlat) 
mapExtentPr <- spTransform(SpatialPoints(mapExtent, 
                  proj4string=CRS("+proj=longlat")),
                  newProj)

can1Pr <- spTransform(Canada, newProj)
plot(mapExtentPr, pch=NA)
plot(can1Pr, border="black", col="lightgrey", add=TRUE)

# loop over the provinces, colouring them according to the number of 
# workshops that have occured there
for (jj in 1:length(colInd)){
  plot(can1Pr[provInd[jj],], border="black", col=colVals[colInd[jj]], add=TRUE)
}
legend('topleft', legend=c(1,NA,10,NA,20,NA,30,NA,40,NA,50), fill=colVals[c(1,seq(5, 50, 5))], bty="n", x.intersp = 2, y.intersp = .5)
```
